Making Deep Q-learning Methods Robust to Time Discretization

### 2019-01-19 paper/icml_drau/advup.tex

To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits a learnable action-dependent limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi_\deltat(s)$ is of order
$\bigO(\deltat)$. This amounts to splitting $Q$ into value and advantage,
and observing that these scale very differently when $\deltat\to 0$.
%Contrary to the action state value function,
This
rescaled advantage function converges when $\deltat\to 0$
to an action-dependent quantity which, if it can be learned, contains the necessary information for policy improvement:

@@

  \begin{theorem}
    There exists $A^\pi(s, a)$ such that for all $a, s$:
    \begin{equation}
      A^\pi_\deltat(s, a) \rightarrow_{\deltat \rightarrow 0} A^\pi(s, a)
    \end{equation}
    Moreover, $A^\pi$ contains all the necessary information for policy improvement. The policy $\pi$ optimal if and only if $A^\pi(s, \pi(s)) = \max_{a'}A^\pi(s, a')$ for all $s \in {\cal S}$.
    \TODO{Is this true? What is true? Is this real life? We should probably add a reasonable hypothesis, for example $\pi_\deltat^* \rightarrow \pi^*$}
  \end{theorem}

The discretized $Q$-function rewrites as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way to approximate $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or $Q$-learning to a reparameterized $Q$-function approximator
\begin{equation}
	Q_\Theta(s, a) \deq V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta \deq (\theta, \psi)$. (To avoid cumbersome notation, 
parameter indices are dropped in what follows.)  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independantly of $\deltat$, this parameterization
provides reasonable scaling of the contribution of actions versus states
in $Q$.
Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
Still, this reparameterization does not, on its own, guarantee that
when $Q$ correctly approximates $Q^\pi_\deltat$, $A$ approximates $A^\pi_\deltat$\footnote{
	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is
	the same. \TODO{still unclear why we'd care about identifiability}
}.
Indeed, for any given pair $(V,\,A)$, the pair $(V(s) -
f(s),\,A(s,a)+f(s)/\deltat)$ (for an arbitrary $f$)
% \begin{align}
% 	\tilde{V}(s) &= V(s) - f(s),\qquad
% 	\tilde{A}(s, a) = A(s, a) + \frac{f(s)}{\deltat}
% \end{align}
yields the exact same $Q$ function. Consequently, $Q$ can correctly approximate
$Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A =
A^\pi_\deltat$ and $V = V^\pi_\deltat$: indeed
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) \deq \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi_\deltat$,
from which a proper ranking of actions can be derived for any $\deltat$.

>>>>

To define an object which contains the same information on actions as
$Q^\pi_\deltat$, but admits a learnable action-dependent limit, it is
natural to define \cite{adv_upd}
\begin{align}
	A^\pi_\deltat(s, a) &\mathdef \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat},
    \label{eq:adv}
\end{align}
a rescaled version of the advantage function, as the difference between between
$Q^\pi_\deltat(s, a)$ and $V^\pi_\deltat(s)$ is of order
$\bigO(\deltat)$. This amounts to splitting $Q$ into value and advantage,
and observing that these scale very differently when $\deltat\to 0$.

Contrary to the action state value function,
this
rescaled advantage function converges when $\deltat\to 0$
to an action-dependent quantity which, if it can be learned, contains the
necessary information for policy improvement \TODO{See Thm... in the appendix; we
should put the theorem in the appendix, because otherwise the theorem is
just a paraphrase of the previous sentence}.

@@

%   \begin{theorem}
%     There exists $A^\pi(s, a)$ such that for all $a, s$:
%     \begin{equation}
%       A^\pi_\deltat(s, a) \rightarrow_{\deltat \rightarrow 0} A^\pi(s, a)
%     \end{equation}
%     Moreover, $A^\pi$ contains all the necessary information for policy improvement. The policy $\pi$ optimal if and only if $A^\pi(s, \pi(s)) = \max_{a'}A^\pi(s, a')$ for all $s \in {\cal S}$.
%     \TODO{Is this true? What is true? Is this real life? We should probably add a reasonable hypothesis, for example $\pi_\deltat^* \rightarrow \pi^*$}
%   \end{theorem}

The discretized $Q$-function rewrites as
\begin{equation}
	Q^\pi_\deltat(s, a) = V^\pi_\deltat(s) + \deltat A^\pi_\deltat(s, a).
	\label{eq:reparam_q_pi}
\end{equation}
A natural way to approximate $V^\pi_\deltat$ and $A^\pi_\deltat$ is to apply
\emph{Sarsa} or $Q$-learning to a reparameterized $Q$-function approximator
\begin{equation}
	Q_\Theta(s, a) \deq V_{\theta}(s) + \deltat A_{\psi}(s, a).
\end{equation}
with $\Theta \deq (\theta, \psi)$. (To avoid cumbersome notation, 
parameter indices are dropped in what follows.)  At initialization, if both $V_{\theta}$ and
$A_{\psi}$ are initialized independently of $\deltat$, this parameterization
provides reasonable scaling of the contribution of actions versus states
in $Q$.
Our goal is for $V_\theta$ to approximate $V^\pi_\deltat$ and for
$A_{\psi}$ to approximate $A^\pi_\deltat$.

%\TODO{CT and LB are fighting each other about the importance which should be given to the next paragraph. A impartial judge should give his opinion. (CT is wrong)}
Still, this reparameterization does not, on its own, guarantee that $A$
correctly approximates $A^\pi_\deltat$ if
$Q$ approximates $Q^\pi_\deltat$% \footnote{
% 	Ensuring that $A$ does not differ from $A^\pi_\deltat$ by a state dependent function only affects
% 	$A$'s interpretability. Indeed, for any function $f$, the ranking of actions given by
% 	$A^\pi_\deltat(s, \cdot)$ and $A^\pi_\deltat(s, \cdot) + f(s)$ is
% 	the same. \TODO{still unclear why we'd care about identifiability}
%}
.
Indeed, for any given pair $(V,\,A)$, the pair $(V(s) -
f(s),\,A(s,a)+f(s)/\deltat)$ (for an arbitrary $f$)
% \begin{align}
% 	\tilde{V}(s) &= V(s) - f(s),\qquad
% 	\tilde{A}(s, a) = A(s, a) + \frac{f(s)}{\deltat}
% \end{align}
yields the exact same $Q$ function. This new $A$ still defines %the same $Q$ and
the same ranking of actions, yet this phenomenon might cause
numerical problems or instability of $A$ when $\deltat\to 0$, and prevents direct
interpretation of the learned $A$.
% Consequently, $Q$ can correctly approximate
% $Q^\pi$ even if $A$ is arbitrarily far from $A^\pi$.
To enforce identifiability of $A$, one must enforce the consistency equation
\begin{equation}
	Q^\pi_\deltat(s, \pi(s)) - V^\pi_\deltat(s) = 0
\end{equation}
on the approximate $A$ and $V$. This translates to
\begin{equation}
	A(s, \pi(s)) = 0.
\end{equation}
With this additional constraint, if $Q = Q^\pi_\deltat$, then $A =
A^\pi_\deltat$ and $V = V^\pi_\deltat$: indeed
\begin{align}
	A^\pi_\deltat(s, a) &= \frac{Q^\pi_\deltat(s,a) - V^\pi_\deltat(s)}{\deltat}\\
		    &= \frac{Q(s,a) - Q(s, \pi(s))}{\deltat}\\
		    &= A(s, a).
\end{align}
In the spirit of~\cite{dueling_nets}, this condition is enforced by writing $A$ as
\begin{equation}
\label{eq:Aparam}
	A(s, a) \deq \bar{A}(s, a) - \bar{A}(s, \pi(s)).
\end{equation}
With this parameterization, one can hope to learn an approximation of $A^\pi$,
from which a proper ranking of actions can be derived for any $\deltat$.
